---
title: "lab11 Assignment"
author: "Somnath Mukherjee"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###  Lab set-up 
```{r,warning=FALSE,message=FALSE,echo=FALSE}
#' a function to only install needed but unavailable packages
#' and loads these packages after installation
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
if (length(new.pkg))
install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}
# use function to install and load packages
packages <- c("ggplot2","MASS","aod","ResourceSelection","lme4","lmerTest","multcomp","gmodels","car","lsmeans","here")
ipak(packages)
```

### Read input file
```{r}
nested = read.table(here('lab11','data',"data11.txt"), header = TRUE)
nested$treatment = as.factor(nested$treatment)
gas = read.table(here('lab11','data',"data11_set2.txt"), header = F)
colnames(gas)=c("OBS","ADDITIVE","RUN","MPG","CAR")
gas$ADDITIVE=as.factor(gas$ADDITIVE)

```

## Part I. Answer the following questions for data set of nested design:  
### Q1.According to the model, do different treatments have significant effect on the apple weight?     

```{r}
# To fit a nested model
fit <- lmer(weight ~ treatment +(1|tree/treatment),data=nested)
```


```{r}
anova(fit)
```
  
Yes, since the p-value is significant (< 0.05), we can conclude that different treatments have significant effect on the apple weight.    


### Q2.Using the contrast to test the hypothesis that the difference between treatment 1 and the average of treatment 2 and treatment 3 is equal to zero. What is your conclusion?   
```{r}
#post-hoc analysis to be able to do Tukey Contrasts amongst the three different treatments:
posthoc = glht(fit,linfct = mcp(treatment="Tukey"))
mcs = summary(posthoc,test=adjusted("single-step"))
mcs
```

Since the p-value is greater than 0.05, we fail to reject the null and conclude that the difference between treatment 1 and 2 is not significant different.  
Since the p-value is less than 0.05, we reject the null and conclude that the difference between treatment 1 & 3 and treatment 2 & 3 is significantly different.    

```{r}
#Tukey grouping
cld(mcs,level=0.05,decreasing=TRUE)
```

From the above tukey grouping result, we can find that treatment 1 and 2 is not significantly different because they are in the same group **b**.  
The difference between treatment 1 and 3 is significant different because they do not belong in the same group. (treatment 1 is in group **b** and treatment 3 is in group **a**)  
The difference between treatment 2 and 3 is significant different because they do not belong in the same group. (treatment 2 is in group **b** and treatment 3 is in group **a**)  


## Part II. Answer the following questions for data set of random block design:  
### Q1.Take the variable Car as a random block. Do different additives have significantly different effect on boosting gas mileage?      


```{r}
#create a mixed model
fit3 <- lmer(MPG ~ ADDITIVE + (1 | CAR)+(1|CAR:ADDITIVE), data = gas)
summary(fit3)
```


```{r}
anova(fit3)
```

Since the p-value is greater that 0.05, we fail to reject null and conclude that different additives do not have significantly different effect on boosting gas mileage.    

### Q2.Take the variable Car as a fixed effect. Do different additives have significantly different effect on boosting gas mileage?      
   

```{r}
#create a factorial design
fit2 = lm(MPG ~ ADDITIVE + CAR + ADDITIVE:CAR, data = gas)
Anova(fit2)
```

Yes, since the p-value is less than 0.05, we can conclude that the different additives have significantly different effect on boosting gas mileage.  

### Q3.Are your answers for question 1 and 2 consistent? If not, why (that is, why the test statistic and p-value are different)?  
No, they are not consistent becuase the formula used to calculate the test-statistics are different and they are fundamentally two different kind of test. Hence, the resultant p-values are different.      
For the case-1 we have random effect(we are comparing variance, actually we are interested in all car-types for different additives which are the our interest as traetment), whereas for the case-2 we have fixed effect (we are comparing the means for specific car-type A,B,D).    

### Q4. Which model is more appropriate? Why?
Random effect model is more appropriate. The reason is that we don’t consider the the difference between these specific 4 car-types, rather all car-types for different additives.        

### Q5. What is the additional assumption we must have for the RBD model? Can we test it?  
The block should be independent. No, we can not test the independence.    
