---
title: "EXST 7014: Lab1 Assignment"
author: "Somnath Mukherjee"
date: "January 22, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###  Install and load required packages

```{r,echo=FALSE,message=FALSE,warning=FALSE}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("knitr", "kableExtra", "ggplot2","plotly")
ipak(packages)

```


###  Read the input file before answering questions

```{r}
fw07p06 = read.table("data_lab1.txt", header = FALSE, sep = "", dec = ".")
colnames(fw07p06)=c("CITY","STATE","LAT","RANGE")
```



### 1. Produce a scatter plot to show the relationship between RANGE and LAT. What is your observation?

```{r}
ggplot(data = fw07p06,aes(x = LAT,y = RANGE))+
  geom_point()+
  theme_classic()

```


There seems to be a positive correlation between LAT (explanatory) and RANGE (dependent) variables.The relationship looks weak with some scatters. 
We can also observe a few potential outliers.  

###   2. Use lm to fit the linear model: y = b0 + b1x + e. Explain briefly your findings,which should include the parameter estimates, the interpretation of the parameters and appropriate hypothesis test.


```{r}
## Fitting the model
model1 <- lm(RANGE ~ LAT, data=fw07p06)
summary(model1)
```


Based on the R output above the regression coefficients are intercept(b0) = -6.47 and slope(b1) = 0.7515,which implies for 1 unit increase in LAT there will be 0.7515 increase for RANGE, and when LAT is zero RANGE will have a value of -6.47.    


Hypotheses test: (We are assuming a significance level of 0.05)   
Null: b1 = 0   
Alternate: b1 != 0   
In other words we assume slope b1 in the below formula is zero.    
                   RANGE(predicted) = b0 + b1*LAT  
Which essentially means LAT deos not help to predict RANGE in our model.              
As we get a very small P-value(4.79e-06) < 0.05 from our test statistics, we will reject our null hypothese.  
Hence we can probably say LAT is a good predictor of RANGE and our model might be a good fit. How good? - that is yet to be investigated.      

###   3. Write the estimated regression function.


```{r}
model1$coefficients

```

Hence our model becomes:  
RANGE(predicted) = -6.47 + 0.7515*LAT.  
The regreession model plotted with ggplot2. 

```{r}
ggplot(data = fw07p06,aes(x = LAT,y = RANGE))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
```



###   4. Find the confidence interval for the intercept and the slope coefficients.  

```{r}
confint(model1) #with default 5% significance level.
```


### 5. What proportion of the variability in the dependent variable RANGE is accounted for by LAT through the regression line ?  

```{r}
summary(model1)$r.squared 
```

The explanatory variable LAT explains approximately 38.85% of the variability for the dependent RANGE variable, through the regression line.
