---
title: "lab12 Assignment"
author: "Somnath Mukherjee"
date: "April 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=3)
```



```{r include=FALSE}
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
if (length(new.pkg))
install.packages(new.pkg, dependencies = TRUE)
sapply(pkg, require, character.only = TRUE)
}
# use function to install and load packages
packages <- c('lme4', 'EMSaov')
ipak(packages)

#lme4 # to fit the linear mixed-effect model
#EMSaov # to extract expected mean square(EMS) formulae for the ANOVA components
```



```{r include=FALSE}
#' Create an object called @auto to host the data set
#'
#' @colClasses= to specify the respective data classes of each column
#' N/B the 1st 3 columns are factor sor repeat (@rep) "factor" 3 times
#' the last column is an integer so, "integer"
auto <- read.table(header = T,
colClasses = c(rep("factor", 3), "integer"),
text=' car driver additive reduct
1 I A 21
2 I B 26
3 I D 20
4 I C 25
1 II D 23
2 II C 26
3 II A 20
4 II B 27
1 III B 15
2 III D 13
3 III C 16
4 III A 16
1 IV C 17
2 IV A 15
3 IV B 20
4 IV D 20 ')
str(auto)
```


```{r include=FALSE}
#Get exhaustive notes with examples on fitting mixed effects models in R; for Linear Mixed and multilevel modelling -from the lme4 package.
vignette(topic="lmer", package = "lme4")
```

### Q1. According to the model, does the additive type have any significant effect on reducing NOX emission? Can the effects of Cars or Drivers be ignored in this model?

```{r}
# Standard Model fitted with Restricted (Residual) Maximum Likelihood Estimation
#The model object is named to mirror the different models fitted using SAS
#random effects are specified with the pipe symbol sandwiched between 1 and the name of the random effect variable e.g. (1|car)
fit.REML <- lmer(reduct ~ additive + (1|car) + (1|driver), data=auto)
summary(fit.REML)
anova(fit.REML) # Get Type 3 Anova Test
```


```{r }
#Model with EMS Details
#"additive" is treated as a fixed effect,
#"car" and "driver" are treated as random effects.
#Therefore, @type = c("F","R","R") following their order in the formula
fit.Type3 <- EMSanova(reduct ~ additive+car+driver, data=auto,
type=c("F","R","R"))
#Remove or Pool interaction (:) effects to the error/residual term
del.ID <- c("additive:car", "additive:driver", "car:driver", "Residuals")
fitType3_Pooled <- PooledANOVA(fit.Type3, del.ID)
fitType3_Pooled
```

(1) Yes, since the p-value is less than 0.05, we can conclude that the additive type have any significant effect on reducing NOX emission.  

(2) Since the p-value is greater than 0.05, we can conclude that the effects of cars can be ignored in this model.  

In addition, since the p-value is less than 0.05, we can conclude that the effects of drivers can not be ignored in this model.  


```{r include=FALSE}
#' Get quick help with examples on the EMSanova and Pooled ANOVA functions
?EMSanova
?PooledANOVA
```

```{r include=FALSE}
#Checking Assumptions
#Test for Homogeneity of Variance
bartlett.test(reduct ~ additive, data=auto)
```

```{r include=FALSE}
#Testing for Normality
#' (Residuals vs. Fitted Values) and QQ Plots colored by Additive types
plot(fit.REML, col=auto$additive, which=c(1,2))
#' Q-Q Plot with Confidence Bands
require(MASS)
library("car")
qqp(residuals(fit.REML), "norm")
#' Alternative QQ-Plot
qqnorm(residuals(fit.REML))
qqline(residuals(fit.REML))
```

```{r include=FALSE}
#' Test for Normality of Residuals
shapiro.test(residuals(fit.REML))
```


### Q2. There is an important assumption for Latin Square Design besides the usual assumptions for the linear model (normality, homogeneous variance et al). What is that?

The Latin Square Design is appropriate only if effects of all three factors (row block, column block and
treatment) are additive, i.e., all interactions are zero.  



