---
title: "Lab3 Assignment"
author: "Somnath Mukherjee"
date: "Feb 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###  Lab set-up 

```{r,warning=FALSE,message=FALSE,echo=FALSE}
#load required packages
#install.packages('olsrr')
#install.packages('here')
#install.packages('car')
library(car)
library(olsrr) #package that runs residual plots and check assumptions for normality
library(here)
library(ggplot2)
```


```{r}
#read input file
theData = read.table(here('lab3','data','data3.txt'), header = F, stringsAsFactors = T) #use here() function to specify path
colnames(theData)=c("CITY","STATE","LAT","RANGE")
```

The latitude (LAT) and the mean monthly range (RANGE), which is the difference between
mean monthly maximum and minimum temperatures, are given for a selected set of US
cities. The following program performs a SLR using RANGE as the dependent variable and
LAT as the independent variable.      
###  Creating a Scatter Plot first  

```{r}
ggplot(data = theData,aes(x = LAT,y = RANGE))+
  geom_point()+
  ggtitle('Scatterplot of Temperature versus Latitude') +
  theme_classic()

```

###  Fitting the SLR model

Based on the scatterplot produced above, we assume that an appriopriate regression model
relating RANGE and LAT is the linear model given by  
  y = b0 + b1x + e  
where y is the RANGE, x is the LAT, and e is a random error term that is normally distributed with the mean 0 and the unknown variance.  
b0 is the estimate of the y-intercept and  b1 is the estimate of the slope coeffient.  

  
```{r}
#fit the model
model1 <‐ lm(RANGE ~ LAT, data = theData)
summary(model1)

```
The regression model is:    
RANGE(predicted) = -6.47 + 0.751*LAT  

### Plot the regression line/model  
```{r}
ggplot(data = theData,aes(x = LAT,y = RANGE))+
  geom_point()+
  geom_smooth(method = lm)+
  ggtitle('SLR of Temperature versus Latitude') +
  theme_classic()
```

The model does not fit the data very well as we can see in the plot above there are a good number of regression outliers.  

### 1. Is the normality assumption violated? State the name and the value of the statistic that you used to reach your conclusion.  

```{r}
ols_test_normality(model1)
```

From Shapiro-Wilk test p-value is < 0.05.Hence, the normality of the residual in not preserved.  
Below is the distribution of fitted-vs-residual plot, which appearently shows non-homogenity.    

```{r}
ols_plot_resid_fit(model1)
```

### 2. Does there appear to be any possible influential observation. State the name and the value of the statistics that you used to reach your conclusion.  

Below will list  Hat Diag H, dffits and dfbetas. It also has cook’s d bar and cov.r. These statistics are usually used to detect possible outliers.  
Astericks beside obs. denotes influential obs.  

```{r}
# Get ALL INFLUENCE MEASURES DISCUSSED (DFFITS,DFBETAS,COOKSD,HAT diag)
influence.measures(model1)
```

Now going by each one of the measures separately:  

*a) Cook’s d chart and analysis*        

```{r}
CkDchart=ols_plot_cooksd_chart(model1) # Chart
```


```{r}
CkDchart$outliers 
```

Above contains the outliers computed by this method.  

*b) To create the DFBETAs Panel we again ask ols to apply dfbetas analysis on our model as follows:*    
```{r}
Dfb = ols_plot_dfbetas(model1) #dfbetas analysis
Dfb$outlier
```

*c) To compute the Dffits one should use:*  
```{r}
Dff=ols_plot_dffits(model1) #dffites analysis
Dff$outlier
```


*d) For the studentized Residual Plot the command is*  
```{r}
StRes=ols_plot_resid_stud(model1) #Studentized Residual analysis - rstudent
StRes$outlier
```

*e) For the Standardized Residual Chart you can use*  
```{r}
ResSta=ols_plot_resid_stand(model1)
ResSta$outlier
```

### 3. What is the confidence interval for the mean Range for all cities with latitude 32.3?

```{r}
predict(model1,data.frame(LAT=32.3),interval="confidence") #confidence for mean at 95%

```

### 4. What is the confidence interval for Range for a randomly selected city with latitude 32.3?

```{r}
predict(model1,data.frame(LAT=32.3),interval="prediction") #confidence for fitted at 95%
```
