---
title: "lab3 Classwork"
author: "Somnath Mukherjee"
date: "January 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###  Lab set-up 

```{r,warning=FALSE,message=FALSE}
#load required packages
#install.packages('olsrr')
#install.packages('here')
#install.packages('car')
library(car)
library(olsrr) #package that runs residual plots and check assumptions for normality
library(here)
library(ggplot2)
```


```{r}
#read input file
theData = read.table(here('lab2','data','data2.txt'), header = F, stringsAsFactors = T) #use here() function to specify path
colnames(theData)=c("CITY","STATE","LAT","RANGE")
```

The latitude (LAT) and the mean monthly range (RANGE), which is the difference between
mean monthly maximum and minimum temperatures, are given for a selected set of US
cities. The following program performs a SLR using RANGE as the dependent variable and
LAT as the independent variable.  

```{r}
ggplot(data = theData,aes(x = LAT,y = RANGE))+
  geom_point()+
  ggtitle('Scatterplot of Temperature versus Latitude') +
  theme_classic()

```

###  Fitting the SLR model

Based on the scatterplot produced above, we assume that an appriopriate regression model
relating RANGE and LAT is the linear model given by  
  Y = b0 + b1X + e  
where Y is the RANGE, X is the LAT, and e is a random error term that is normally distributed with the mean 0 and the unknown variance.  
b0 is the estimate of the Y-intercept and  b1 is the estimate of the slope coeffient.  

  
```{r}
#fit the model
SLR_model <‐ lm(RANGE ~ LAT, data = theData)
summary(SLR_model)

```

### Observation Diagonistics[Regression outlier, Leverage obs.,influential obs.]  
**Regression outlier:** We use rstudent(Deleted Studentized residual) measure to check for regression outlier.  
For our dataset if the value is >2 or <-2 then the point is an outlier.

```{r}
## R Student - used to check the regression outlier.
rStudent <‐ rstudent(SLR_model)
rStudent
```


```{r}
outlierTest(SLR_model) # run test to get possible outliers considering p-value using adjusted alpha level.(part of car package)

```

**Leverage observation:** Unusally large X value than mean X.  

```{r}
## Hat diagonal values - to identify leverage outlier(unusual X value)
HatDiag <‐ lm.influence(SLR_model)$hat # get Hat Diag values
cutoff <‐ 2*(length(coef(SLR_model))/length(HatDiag)) # cu‐off at 2*p/n
a <‐ theData[which(lm.influence(SLR_model)$hat > cutoff),] # Print obs with HAT > 2*p/n
cbind(a, HatDiag[HatDiag > cutoff]) #obs with high hatdiag value
plot(HatDiag, ylab="HatDiag")
```

**Influential observation** (based on DFFITS,DFBETAS,COOKSD):  
Astericks beside obs. denotes influential obs.

```{r}
# Get ALL INFLUENCE MEASURES DISCUSSED (DFFITS,DFBETAS,COOKSD)
influence.measures(SLR_model)
```

```{r}
## To get individual outliers(influential obs.) run the codes below.
## based on DFFITS
DFFITS_model <‐ abs(dffits(SLR_model)) # get absolute values of DFFITS
b <‐ theData[which(DFFITS_model > 1),] # get DFFITS > 1
cbind(b, DFFITS = DFFITS_model[DFFITS_model > 1]) # Print obs with DFFITS >1
plot(dffits(SLR_model), ylab="DFFITS") # Plot all DFFITS
```


```{r}
## DFBETAS (we are only checking for b1[LAT] for our case but b0[intercept] is also valid check)
DFBETAS_model <‐ abs(dfbetas(SLR_model)[, 'LAT']) # get absolute values of DFBETAS for LAT
c <‐ theData[which(DFBETAS_model > 1),] # get LAT DFBETAS > 1
cbind(c, DFBETA_LAT = DFBETAS_model[DFBETAS_model > 1]) # Print obs with LAT DFBETAS > 1
plot(DFBETAS_model, ylab="DFBETA (LAT)") # Plot LAT DFBETAS
```


```{r}
## COOK's Distance
COOKS_mod <‐ cooks.distance(SLR_model) # get cook's D for all observations
cutoff <‐ 4/length(COOKS_mod) # cut off at 4/n
d <‐ theData[which(COOKS_mod > cutoff),]
cbind(d, COOKsD = COOKS_mod[COOKS_mod > cutoff])


```

2. Evaluate the assumptions of SLR using Residual Plots and the Normality Test.  

```{r}
ols_plot_resid_fit(SLR_model)
ols_plot_resid_qq(SLR_model)
ols_test_normality(SLR_model) ## Test normality
```

